An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-04-22 EconGrowth.Rnw
Edited the text. Switched around the LINEX function to put the exponential term first.

@@

\item{\st{Heun: Investigate use of the micEcon package transLogEst function for CD parameter estimation. Doesn't look like it will work, because it doesn't do the CD form of interest to us.}}

@@

\item \st{Heun: Ask Martin if he can supply an ``Economics'' affiliation. Will probably make a stronger paper.}
\item \st{Heun: Create a table of AIC values for each fit.}
\item \st{Heun: Add $u$ predictions for Cobb-Douglas}
\item \st{Heun: Move tables to an appendix at the end.}
\item \st{Heun: Add acknowledgements (Dad, Pruim)}
\item \st{Heun: Add $u$ parameter table and graph}
\item \st{Heun: Add covariance metrics.}
\item \st{Heun: Add single factor fits}

@@

The Cobb-Douglas with energy parameter estimation process is challenging. To demonstrate the characteristics of the parameter estimation process, we performed a grid search over values of $\gamma$, wherein we re-estimated values of $\lambda$, $\alpha$, and $\beta$ at each value of $\gamma$. Figure \ref{fig:CDeGammaGraphsZA} shows the sum of squared errors ($SSE$), $\alpha$, and $\beta$ as a function of the fixed value of $\gamma$ for South Africa with thermal energy ($q$). The minimum $SSE$ occurs at a discontinuity caused by the constraint $\beta \geq 0.0$. The parameter estimation algorithm has difficulty recognizing the values of $\gamma$, $\alpha$, $\beta$, and $\lambda$ that produce minimum $SSE$, because the constrained model behaves very differently on either side of the discontinuity.

@@

The key energy parameters for the Cobb-Douglas and CES models are $\gamma$, $\delta$, and $\sigma$. The parameter $\gamma$ in the Cobb-Douglas model (Equation \ref{eq:CD_With_Energy}) represents the importance of energy for economic growth. When $\gamma$ is small (close to 0), energy is not a significant contributor to economic growth for an economy. Conversely, when $\gamma$ is large (close to 1), energy is the primary contributor to economic growth. However, if the 95\% confidence interval for $\gamma$ for any given economy extends to 0.0, we cannot say with high confidence that energy makes a meaningful contribution to economic growth for that economy. On a graph of the width of the 95\% confidence interval for 95\% confidence interval on $\gamma$ ($CI_{95\%,\gamma}$) vs. $\gamma$, where each economy is represented by a point, the region where the 95\% confidence interval on $\gamma$ crosses 0 is an upper-left triangle, above and to the left of a line that connects points (0, 0) and (1, 2) on the graph. If $\gamma$ for a given economy is below and to the right of the line connecting (0, 0) and (1, 2), we have high confidence that we can say whether or not energy is a significant contributor to economic growth. Thus, only when $\gamma$ is below the line and small (i.e., close to 0) for an economy, can we say with confidence that energy is an insignificant contributor to economic growth. Only when $\gamma$ is below the line and large (i.e., close to 1) for an economy can we say that energy is a significant contributor to economic growth, according to the Cobb-Douglas model.

@@

For the CES model (Equation \ref{eq:micEconCES_model}), the parameter $\delta$ indicates that importance of energy for economic growth. When $\delta$ is close to 1, energy is an insignificant contributor to economic growth, and when $\delta$ is close to 0, energy is the primary contributor to economic growth for an economy. However, if the 95\% confidence interval for $\delta$ extends to 1, we cannot say with high confidence that energy makes a meaningful contribution to economic growth for that economy. On a graph of $\delta$ confidence interval width ($CI_{95\%,\delta}$) vs. $\delta$, where each economy is represented by a point, the region where the 95\% confidence interval for $\delta$ crosses 1 is an upper-right triangle, above and to the right of a line that connects points (1, 0) and (0, 2) on the graph. If $\delta$ for a given economy is below and to the left of the line connecting (1, 0) and (0, 2), we have high confidence that we can say whether or not energy is a significant contributor to economic growth. Thus, only when $\delta$ is below the line and large (i.e., close to 1) for an economy, can we say with confidence that energy is an insignificant contributor to economic growth. Only when $\delta$ is below the line and small (i.e., close to 0) for an economy can we say that energy is a significant contributor to economic growth, according to the CES model.

@@

\begin{equation} \label{eq:LINEX_model}
  y = e \: \mathrm{e}^{\left[a_0 \left(2 - \frac{l + e}{k} \right) + a_0 c_t \left(\frac{l}{e} - 1 \right)\right]} ,
\end{equation}

>>>>

\item{Heun: Add diagonal lines to triangle plots via the doubleY graph idea. Make a simple data.frame that contains the coordinates of the line and add it to the graph as a second Y axis, but don't show the second Y axis.}

@@

\item{\st{Heun: Investigate use of the micEcon package transLogEst function for CD parameter estimation. Result: doesn't look like it will work, because it doesn't do the CD form of interest to us.}}

@@

\item{\st{Heun: Ask Martin if he can supply an ``Economics'' affiliation. Will probably make a stronger paper.}}
\item{\st{Heun: Create a table of AIC values for each fit.}}
\item{\st{Heun: Add $u$ predictions for Cobb-Douglas}}
\item{\st{Heun: Move tables to an appendix at the end.}}
\item{\st{Heun: Add acknowledgements (Dad, Pruim)}}
\item{\st{Heun: Add $u$ parameter table and graph}}
\item{\st{Heun: Add covariance metrics.}}
\item{\st{Heun: Add single factor fits}}

@@

The Cobb-Douglas with energy parameter estimation process is challenging. To demonstrate the characteristics of the parameter estimation process, we performed a grid search over values of $\gamma$, wherein we re-estimated values of $\lambda$, $\alpha$, and $\beta$ at each value of $\gamma$. Figure \ref{fig:CDeGridLattice} shows the sum of squared errors ($SSE$), $\alpha$, and $\beta$ as a function of the fixed value of $\gamma$ with thermal energy ($q$). For most economies, the minimum $SSE$ occurs at a boundary of $\alpha$, $\beta$, or $\gamma$. The parameter estimation algorithm has difficulty assessing confidence intervals for $\alpha$, $\beta$, and $\gamma$, because the constrained model behaves very differently on either side of the constraint.

@@

The key energy parameters for the Cobb-Douglas and CES models are $\gamma$, $\delta$, and $\sigma$. The parameter $\gamma$ in the Cobb-Douglas model (Equation \ref{eq:CD_With_Energy}) represents the importance of energy for economic growth. When $\gamma \rightarrow 0$, energy is not a significant contributor to economic growth for an economy. Conversely, when $\gamma \rightarrow 1$, energy is the primary contributor to economic growth. However, if the 95\% confidence interval for $\gamma$ for any given economy extends to 0.0, we cannot say with high confidence that energy makes a meaningful contribution to economic growth for that economy. On a graph of the width of the 95\% confidence interval for 95\% confidence interval on $\gamma$ ($CI_{95\%,\gamma}$) vs. $\gamma$, where each economy is represented by a point, the region where the 95\% confidence interval on $\gamma$ crosses 0 is an upper-left triangle, above and to the left of a line that connects points (0, 0) and (1, 2) on the graph. If $\gamma$ for a given economy is below and to the right of the line connecting (0, 0) and (1, 2), we have high confidence that we can say whether or not energy is a significant contributor to economic growth. Thus, only when $\gamma$ is below the line and small (i.e., close to 0) for an economy, can we say with confidence that energy is an insignificant contributor to economic growth. Only when $\gamma$ is below the line and large (i.e., close to 1) for an economy can we say that energy is a significant contributor to economic growth, according to the Cobb-Douglas model.

@@

For the CES model (Equation \ref{eq:micEconCES_model}), the parameter $\delta$ indicates that importance of energy for economic growth. When $\delta \rightarrow 1$, energy is an insignificant contributor to economic growth, and when $\delta \rightarrow 0$, energy is the primary contributor to economic growth for an economy. However, if the 95\% confidence interval for $\delta$ extends to 1, we cannot say with high confidence that energy makes a meaningful contribution to economic growth for that economy. On a graph of $\delta$ confidence interval width ($CI_{95\%,\delta}$) vs. $\delta$, where each economy is represented by a point, the region where the 95\% confidence interval for $\delta$ crosses 1 is an upper-right triangle, above and to the right of a line that connects points (1, 0) and (0, 2) on the graph. If $\delta$ for a given economy is below and to the left of the line connecting (1, 0) and (0, 2), we have high confidence that we can say whether or not energy is a significant contributor to economic growth. Thus, only when $\delta$ is below the line and large (i.e., close to 1) for an economy, can we say with confidence that energy is an insignificant contributor to economic growth. Only when $\delta$ is below the line and small (i.e., close to 0) for an economy can we say that energy is a significant contributor to economic growth, according to the CES model.

@@

\begin{equation} \label{eq:LINEX_model}
  y = \mathrm{e}^{\left[a_0 \left(2 - \frac{l + e}{k} \right) + a_0 c_t \left(\frac{l}{e} - 1 \right)\right]} \: e ,
\end{equation}
